\documentclass[./main.tex]{subfiles}
\begin{document}

\section{HC Application Tags}

\textbf{NOTE: these are rough tags which indicate where the HC is applied. I will invest effort in fleshing these out over the coming 3 weeks.}

\paragraph*{\textbf{\#estimation:}}\label{hc:estimation} Applied to estimating the relative advantage of compilation vs dynamic execution using the Big-O complexity approximation. Documented in Chapter \ref{chap:macimplementation}.

\paragraph*{\textbf{\#algorithms:}}\label{hc:algorithms} Applied in the construction of the DGP sampling procedure documented in Chapter \ref{chap:macimplementation}.

\paragraph*{\textbf{\#variables:}}\label{hc:variables} Applies extensively in constructing the statistical framework that is used to formalize the concept of a Data Generating Process. See Chapter \ref{chap:problemspace}.

\paragraph*{\textbf{\#sampling:}}\label{hc:sampling} used in construction benchmarks that average over all relevant sources of variance.

\paragraph*{\textbf{\#descriptivestats:}}\label{hc:descriptivestats} used to calculate performance and data metrics from sampled datasets.

\paragraph*{\textbf{\#distributions:}}\label{hc:distributions} used extensively in the theoretical design of maccabee (Chapter \ref{chap:macdesign}) and the actual implementation.

\paragraph*{\textbf{\#probability:}}\label{hc:probability} used extensively in the theoretical foundations in Chapter \ref{chap:framework}.

\paragraph*{\textbf{\#correlation:}}\label{hc:correlation} used extensively in the theoretical foundations in Chapter \ref{chap:framework} and in the analysis in Chapter \ref{chap:macvalidation}.

\paragraph*{\textbf{\#breakitdown:}}\label{hc:breakitdown} pending

\paragraph*{\textbf{\#dataviz:}}\label{hc:dataviz} used in Chapter \ref{chap:macvalidation}.

\paragraph*{\textbf{\#gapanalysis:}}\label{hc:gapanalysis} used in the design of Maccabee (Chap \ref{chap:macdesign}) when frameing the design goals in terms of the weaknesses of existing approaches.

\paragraph*{\textbf{\#modeling:}}\label{hc:modeling} Applies extensively in constructing the statistical framework that is used to formalize the concept of a Data Generating Process. See Chapter \ref{chap:problemspace}.

\paragraph*{\textbf{\#testability:}}\label{hc:testability} used in Chapter \ref{chap:macvalidation}.

\paragraph*{\textbf{\#audience:}}\label{hc:audience} used in the choice to make real online documentation that is easily consumed.

\paragraph*{\textbf{\#communicationdesign:}}\label{hc:communicationdesign} used in the choice to make real online documentation that is easily consumed.

\paragraph*{\textbf{\#critique:}}\label{hc:critique} used extensively in Chapter \ref{chap:litreview} to review existing approaches.

\paragraph*{\textbf{\#strategize:}}\label{hc:strategize} pending
\paragraph*{\textbf{\#selfawareness:}}\label{hc:selfawareness} pending

\paragraph*{\textbf{\#confidenceintervals:}}\label{hc:confidenceintervals} used in Chapter \ref{chap:macvalidation} when discussing benchmark validation.

\paragraph*{\textbf{\#significance:}}\label{hc:significance} used in Chapter \ref{chap:macvalidation} when discussing benchmark validation.

\paragraph*{\textbf{\#biasidentification:}}\label{hc:biasidentification} used in the review of existing methods when discussing how researchers may introduce bias by selecting functions which suit their own methods.
\paragraph*{\textbf{\#biasmitigation:}}\label{hc:biasmitigation} used in the review of existing methods when discussing how researchers may introduce bias by selecting functions which suit their own methods.

\paragraph*{\textbf{\#medium:}}\label{hc:medium} used in the choice to build online docs.
\paragraph*{\textbf{\#context:}}\label{hc:context} used in the lit review in Chapter \ref{chap:litreview}.

\section{LO Application Tags}

\textbf{NOTE: because these are custom LOs, their application is fairly obvious. I will write extensive tags soon.}

\subsection{Foundation: Causal Inference}
\paragraph*{\textbf{\#CausalInferenceTheory:}}\label{lo:CausalInferenceTheory} applied in Chapter \ref{chap:framework}.
\paragraph*{\textbf{\#graphicalmodels:}}\label{lo:graphicalmodels} applied in Chapter \ref{chap:framework}
\paragraph*{\textbf{\#probabilitytheory:}}\label{lo:probabilitytheory} applied in Chapter \ref{chap:framework}
\paragraph*{\textbf{\#CausalProblemSpace:}}\label{lo:CausalProblemSpace} applied in Chapter \ref{chap:problemspace}


\subsection{Hybrid MC Benchmarking: Theory/Design}
\paragraph*{\textbf{\#CausalBenchmarkTheory:}}\label{lo:CausalBenchmarkTheory} applied in Chapter \ref{chap:macdesign}
\paragraph*{\textbf{\#CausalBenchmarkDesign:}}\label{lo:CausalBenchmarkDesign} applied in Chapter \ref{chap:macdesign}

\subsection{Hybrid MC Benchmarking: Implementation}
\paragraph*{\textbf{\#CodeDesign:}}\label{lo:CodeDesign} applied in Chapter \ref{chap:macimplementation}
\paragraph*{\textbf{\#CodeImplementation:}}\label{lo:CodeImplementation} applied in Chapter \ref{chap:macimplementation}
\paragraph*{\textbf{\#CodeUsability:}}\label{lo:CodeUsability} applied in Chapter \ref{chap:macimplementation}

\end{document}